---
title: "Reticulate"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Import reticulate
First we'll import reticulate in an R code block
```{r}
library(reticulate)
```

### Run Python code
We can immediately start playing with python within a code block designated as python
```{python}
import pandas as pd

print('hello python')

some_dict = {'a':1, 'b':2, 'c':3}

print(some_dict.keys())
```


### Import Python modules
With reticulate, we can also import any Python module and call that module from R
```{r}
os <- import("os")
os$listdir(".")

# Notice the code completion for methods available in python modules!!!
sys <- import('sys')
sys$
```

### REPL sessions
We can also start a Python REPL (read-eval-print shell) within our session by calling repl_python() at console (see console).  Notice the message I get when I start REPL on Python version being used
```{r}
repl_python()
# below code at console
# import pandas as pd
# gages = pd.read_csv('https://raw.githubusercontent.com/mhweber/R-User-Group-Spatial-Workshop-2018/gh-pages/data/StreamGages.csv')
# gages.head()
# gages['STATE'].unique()
# PNW_gages = gages[gages['STATE'].isin(['OR','WA','ID'])]
# exit
```

### Access Python objects directly from R
Now work with the pandas data directly within R
```{r}
library(sf)
library(ggplot2)

gages <- st_as_sf(py$PNW_gages,coords = c('LON_SITE','LAT_SITE'),crs = 4269) 
gages <- st_transform(gages, crs=5070) #5070 is Albers system in metres
ggplot(gages) + geom_sf()
```

reticulate converts to equivalent data structures between R and Python
```{r}
class(py$PNW_gages)
```

### Access R objects from Python
Conversely, we can access our R objects within python chunks
```{r}
repl_python()
print(r.gages.head())
# list the column names
list(r.gages)
# what are all the values of STATE?
r.gages.STATE.unique()
```

reticulate converts to equivalent data structures between Python and R 
```{python}
print(type(r.gages))
```

We noted earlier how we access Python functions, methods, data via the $ operator.
And that reticulate automatically transforms objects in Python to equivalent objects in R.
Here's an example from [this Rpub](https://rstudio-pubs-static.s3.amazonaws.com/397064_bd288708515c4b71a6e78dfe91b8bb99.html) that shows where and why we might bypass this.
```{r}
pandas <- import("pandas")
titanic <- pandas$read_csv("https://goo.gl/4Gqsnz")
titanic$describe()
```

We need to override automatic conversion and specify `convert=FALSE`
```{r}
pandas <- import("pandas", convert =FALSE)
titanic <- pandas$read_csv("https://goo.gl/4Gqsnz")
description <- titanic$describe()
```

We can directly convert objects from Python to R, and R to Python with reticulate using `py_to_r()` and `r_to_py`
```{r}
description_r <- py_to_r(description)
description_py = r_to_py(description_r)
```

### Source Python scripts
If we have a python script called 'sum_avg.py' with this code:
```{python}
def sum_avg(x,y):
    return(x+y, (x + y)/2)
```

We can run it in R with:
```{r}
source_python('sum_avg.py')
sum_avg(4,4)
```

### Matrics and numpy arrays
reticulate allows integration between Python numpy arrays and R matrices.  reticulate automatically converts to equivalent R types unless you override (i.e. convert=FALSE).  Treatment of numpy arrays in R deserves consideration!
```{r}
# import numpy into R
np <- import("numpy")
matrix1 <- matrix(data = 1:25, nrow = 5, ncol = 5)
matrix2 <- matrix(data = 25:1, nrow = 5, ncol = 5)

# multiply matrices as numpy arrays
matrix3 <- np$matmul(matrix1, matrix2)

# An R matrix
class(matrix3)

print(matrix3)

# import numpy into R but don't allow default conversion just to contrast
np <- import("numpy", convert=FALSE)
matrix1 <- np$arange(1, 26)$reshape(5L, 5L)
matrix1 <- matrix1$astype(dtype="int") # I wasn't savvy enough to bake this into previous line....
matrix2 <- matrix(data=25:1, nrow=5, ncol=5)
class(matrix1)
class(matrix2)
matrix3 <- np$matmul(matrix1, matrix2)
class(matrix3)
print(matrix3)
```

### Python versions
You can use
- Specific Python binaries you point to
- Python virtual environments
- Conda environments

You can first simply get information on what versions of Python are available and what version of Python and numpy are in use
```{r}
py_config()
```

### Specify python to use 
Either by path or by conda environment:
```{r}
use_condaenv("StreamCat")
use_python("C:/Python27/ArcGISx6410.7")

# Note that to actually work you have to run immediately before you do anything else - once initialized, you can't change Python versions within an R session.
# retiulate
# use_condaenv("StreamCat", required = T)
# use_python("C:/Python27/ArcGISx6410.7", required = T)
```

### Create conda or virtual environment
Create environments, install packages, remove environments in isolated environments
```{r}
conda_create("demo")
conda_install("demo", "pillow")
pil <- import("PIL") # I've added the Python Imaging Library to new 'demo' environment and loaded library into R
conda_remove("demo")

virtualenv_create("demo")
virtualenv_install("demo", "pillow")
pil <- import("PIL") # I've added the Python Imaging Library to new 'demo' environment and loaded library into R
virtualenv_remove("demo")
```